### 2025-05-27 appendices/mpif/gaus.tex
Reverting Chapter 4 image captions, merge appendix from mpif_article

As a reminder, we assume that there are $U\geq 2$ units, labeled $\seq{1}{U}$, with the data for unit $u$ being denoted as $\mathbf{y}^*_u$. 
We write $\theta = (\phi, \psi_1, \ldots, \psi_U)$, and recall that the unit likelihood $L_{u}(\theta;\mathbf{y}^*_u)$ for unit $u \in 1:U$ depends only on the shared parameter $\phi$ and the unit-specific parameter $\psi_u$.
The panel assumption implies that, conditioned on the parameter vector, the units are dynamically independent.
Therefore, we can decompose the likelihood function for the entire collection of data $L(\theta; \mathbf{y}^*)$ as:
\begin{align*}
    L(\theta; \mathbf{y}^*) = \prod_{u = 1}^U L_u(\theta;\mathbf{y}_u^*) = \prod_{u = 1}^U L_u(\phi, \psi_u;\mathbf{y}_u^*).
\end{align*}

@@

We now show that the mean of the resulting Gaussian distribution converges to the MLE, while the covariance matrix converges to the zero matrix, resulting in a density with all mass centered at the MLE. 
To aid this calculation, we introduce the following lemma.
\begin{lemma}\label{lemma:bound}
\label{lemma:matrix}
  Let $d$ be a positive integer, and let $B_k \in \R^{2\times 2}$ for $k \in \seq{1}{d}$ be a collection of real-valued matrices.
  We construct a sequence of matrices $A_{k} \in \R^{d+1\times d+1}$ such that:
  \begin{equation*}
    % \spacingset{1.5}
    [A_k]_{i, j} = \begin{cases}
        [B_k]_{1,1}, & i = j = 1 \\ 
        [B_k]_{1,2}, & i = 1, j = k+1 \\
        [B_k]_{2, 1}, & i = k+1, j=1 \\
        [B_k]_{2,2}, & i = j = k+1 \\
        1, & i = j, \, i \notin \{1, k+1\} \\
        0, & \text{otherwise}
	\end{cases}
  \end{equation*}
 That is, $A_k$ is a perturbation of the identity matrix, where the first and $(k+1)$th row and column have been modified on the diagonal and on their off-diagonal intersection to match the matrix $B_k$.
If for all $k \in \seq{1}{d}$, $\|B_k\|_2 \leq c$ for some constant $0 < \lemmaBound \leq 1$, then
\begin{equation}\label{eq:lemma:bound}
\bigg\| \prod_{k = 1}^d A_k \bigg\|_2 \leq \lemmaBound.
\end{equation}
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lemma:bound}]
  For $i, j \in \seq{1}{(d+1)}$, denote $\mu_{(i, j)} \in \mathbb{R}^2$ as the sub-vector of $\mu \in \mathbb{R}^{d+1}$ that contains only the $i$ and $j$th elements, and write $\mu_{-(i, j)} \in \mathbb{R}^{d-1}$ to be the sub-vector of $\mu$ after removing these elements.
  By design, the matrix $A_k$ operates only on the sub-vector $\mu_{(1, k+1)}$.
That is, if $B_k \, \mu_{(1, k+1)} = (\tilde{\mu}_{(1)}, \tilde{\mu}_{(k+1)})^T$, then $A_k\mu = (\tilde{\mu}_{(1)}, \mu_{(2)}, \ldots, \tilde{\mu}_{(k + 1)}, \ldots, \mu_{(d+1)})$.
  From this, we see that for positive integer $m \leq d$,
  the first $m$ factors of the product $\prod_{k = 1}^d A_k$ modify only the first $m + 1$ dimensions of a vector $\mu \in \R^{d+1}$.

We proceed by mathematical induction on the dimension size.
Let $\big\{A^{(d)}_k, k\in \seq{1}{d-1}\big\}$ be a collection of matrices satisfying the condition of Lemma~\ref{lemma:bound} for each $d$.
Setting $P_d=\big\|\prod_{k = 1}^{d-1} A^{(d)}_k \big\|_2$, we first observe that Eq.~(\ref{eq:lemma:bound}) holds for $d=2$ as a direct consequence of the condition $\|B_k\|_2 \leq c$.
Suppose inductively that the lemma holds for $d$, so that $P_d\leq\lemmaBound$.
We wish to bound $P_{d+1}$.
We can choose $A^{(d)}_k$ to be the $(d+1)\times (d+1)$ sub-matrix of $A^{(d+1)}_k$ omitting row and column $(d+2)$.
Thus, for $k\in\seq{1}{d}$,
\begin{equation*}
A^{(d+1)}_k =
  \begin{pmatrix}
    A^{(d)}_k & 0 \\
    0 & 1
  \end{pmatrix}.
\end{equation*}
Consider a vector $\mu \in \R^{d+1}$, such that $\|\mu\|_2 \leq 1$.
Let $\tilde{\mu} \in \R^{d}$ be defined by 
\begin{equation}
\tilde{\mu} = \bigg[\bigg(\prod_{k = 1}^{d-1} A^{(d+1)}_k\bigg)\mu\bigg]_{(1:d)}
\end{equation}
and notice that we have
\begin{equation}\label{eq:lemma:d}
\tilde{\mu} = \bigg(\prod_{k = 1}^{d-1} A^{(d)}_k\bigg)\mu_{(1:d)}.
\end{equation}
By construction, we have 
\begin{align*}
    \bigg|\Big(\prod_{k = 1}^d A^{(d+1)}_k\Big)\mu \bigg|^2_2 &= \bigg|\Big(A^{(d+1)}_d\prod_{k = 1}^{d-1} A^{(d+1)}_k\Big)\mu \bigg|^2_2\\
    &= \big|A^{(d+1)}_d(\tilde{\mu}_{(1:d)}, \mu_{(d+1)})^T\big|^2_2 \\
    &= \big|B^{(d+1)}_d (\tilde{\mu}_1, \mu_{(d+1)})^T\big|^2_2 + \big|\tilde{\mu}_{(1:d)}\big|^2_2 - \tilde{\mu}^2_1.
\end{align*}
Because $\|B^{(d+1)}_d\|_2 \leq \lemmaBound$, $|B^{(d+1)}_d (\tilde{\mu}_1, \mu_{(d+1)})^T|^2_2 \leq \lemmaBound^2|(\tilde{\mu}_1, \mu_{(d+1)})^T|^2_2$.
Furthermore, our inductive hypothesis applied to Eq.~(\ref{eq:lemma:d}) implies that
$|{\tilde{\mu}}_{(1:d)}|^2_2 \leq \lemmaBound^2|\mu_{(1:d)}|^2_2 \leq \lemmaBound^2|\mu|^2_2$.
Therefore,
\begin{align}
    \bigg|\Big(\prod_{k = 1}^d A_k\Big)\mu\bigg|^2_2 &\leq \lemmaBound^2({\tilde{\mu}}^2_1 + \mu_{(d+1)}^2) + \lemmaBound^2(|\mu_{(1:d)}|^2_2) - {\tilde{\mu}}^2_1 \\
    &= (\lemmaBound^2-1){\tilde{\mu}}^2_1 + \lemmaBound^2 |\mu|^2_2
    \\
    \label{eq:lemma:conclusion}
    &\leq \lemmaBound^2 \|\mu\|^2_2,
\end{align}
with Eq.~(\ref{eq:lemma:conclusion}) implied by $\lemmaBound \leq 1$, and hence $(\lemmaBound^2-1) < 0$.
It follows immediately from Eq.~(\ref{eq:lemma:conclusion}) that $P_{d+1}\le c$, completing the proof.
\end{proof}

We now return to the main argument.
%continue the proof of Theorem~\ref{theorem:GG}.

@@

Performing the Bayes update to this distribution (Eq.~\ref{eq:margBayes}) results in an unmarginalized precision matrix, which we denote as $\tilde{\Gamma}_{\nmif, \unit}$, where the only modified components are
$$
\tilde{\Gamma}^{(\nmif, \unit)}_{\nmif, \unit} = \Gamma^{(\nmif, \unit)}_{\nmif, \unit-1} + \Lambda_\unit^* = \begin{pmatrix}
  \tau^{(\phi)}_{\nmif, \unit-1} + \Lambda^{(\unit)}_{\phi} & \Lambda_{\phi, \unit} \\ 
  \Lambda_{\phi, \unit} & \tau^{(\unit)}_{\nmif, \unit-1} + \Lambda_{\unit}
\end{pmatrix}. 
$$
The corresponding mean $\tilde{\mu}_{(\nmif, \unit)} = \mu_{(\nmif, \unit)}$, noting that the marginalization procedure (Eq.~\ref{eq:MPIFupdate}) does not affect the mean, remains the same as the previous iteration except for the components
\begin{align}
\begin{split}
    \tilde{\mu}^{(\phi, \unit)}_{\nmif, \unit} &= \big(\tilde{\Gamma}^{(\nmif, \unit)}_{\nmif, \unit}\big)^{-1}\big(\Gamma^{(\nmif, \unit)}_{\nmif, \unit-1} \mu_{\nmif, \unit-1}^{(\phi, u)} + \Lambda_u^* (0, 0)^T\big) \\
   \mu^{(\phi, u)}_{\nmif, \unit} &= \big(\tilde{\Gamma}^{(\phi, u)}_{\nmif, \unit}\big)^{-1}\Gamma^{(\phi, u)}_{\nmif, \unit-1} \mu_{\nmif, \unit-1}^{(\phi, u)} \\
   &= B_{\nmif, \unit} \mu_{\nmif, \unit-1}^{(\phi, u)}, 
\end{split}\label{eq:submatB}
\end{align}
where $B_{\nmif, \unit} = \big(\tilde{\Gamma}^{(\phi, 1)}_{\nmif, \unit}\big)^{-1}\Gamma^{(\phi, u)}_{\nmif, \unit-1}$ is a $2\times 2$ matrix that provides the update to the first and $(\unit+1)$th components of a vector $\mu \in \mathbb{R}^{\Unit+1}$.
Now by defining matrix $A_{\nmif, \unit}$ to be a perturbation of the $\Unit+1$ identity matrix, such that
\begin{equation*}
% \spacingset{1.5}
\big[A_{\nmif, \unit}\big]_{i, j} = \begin{cases}
  [B_{\nmif, \unit}]_{1, 1} & i = j = 1 \\
  [B_{\nmif, \unit}]_{2, 2} & i = j = u+1 \\
  [B_{\nmif, \unit}]_{1, 2} & i = 1, j = k+1 \\
  [B_{\nmif, \unit}]_{2, 1} & i = k+1, j = 1 \\
  1 & i = j, i \notin \{1, k+1\} \\
  0 & \text{otherwise}
\end{cases},
\end{equation*}
We see that for a given precision matrix $\Gamma_{\nmif, \unit-1}$, the a single sub-iteration of Eqs.~\ref{eq:margBayes} and \ref{eq:MPIFupdate} corresponds to a linear update of the mean vector $\mu_{\nmif, \unit-1}$, defined by
\begin{align*}
  \mu_{\nmif, \unit} = A_{\nmif, \unit} \, \mu_{\nmif, \unit-1}.
\end{align*}
Thus, the mean after $M$ complete iterations, denoted by $\mu_{M}$ is calculated by the product of these linear transformations
\begin{align*}
  \mu_M = \bigg(\prod_{\nmif = 1}^M\prod_{u = 1}^U A_{\nmif, \unit}\bigg)\mu_0. 
\end{align*}
Thus, the long-term value of $\mu_M$ depends on the long-term behavior of the product of matrices $A_{\nmif, \unit}$.

We now consider the limiting behavior of the precision matrices, which determines the behavior of the matrices $A_{\nmif, \unit}$. 
First, we recall that the unmarginalized precision matrix is of the form:
\begin{align*}
  {\tilde{\Gamma}}^{(\nmif, \unit)}_{\nmif, \unit} = \begin{pmatrix}
  \tau^{(\phi)}_{\nmif, \unit-1} + \Lambda^{(u)}_{\phi} & \Lambda_{\phi, u} \\ 
  \Lambda_{\phi, u} & \tau^{(u)}_{\nmif, \unit-1} + \Lambda_{u}\end{pmatrix} = \begin{pmatrix} 
  a & b \\ b & d
\end{pmatrix}. 
\end{align*}
The marginalization step (Eq.~\ref{eq:MPIFupdate}) for Gaussian densities is represented by setting the off-diagonal elements of the covariance matrix to be zero. 
Thus, 
\begin{align*}
    {\tilde{\Gamma}}^{(\nmif, \unit)}_{\nmif, \unit} &= \begin{pmatrix}a & b \\ b & d \end{pmatrix} \\
    \left({\tilde{\Gamma}}^{(\nmif, \unit)}_{\nmif, \unit}\right)^{-1} &= \frac{1}{ad - b^2}\begin{pmatrix}d & -b \\ -b & a\end{pmatrix} \\
    \left(\Gamma^{(\nmif, \unit)}_{\nmif, \unit}\right)^{-1} &= \frac{1}{ad - b^2}\begin{pmatrix}d & 0 \\ 0 & a\end{pmatrix}\\
    &= \begin{pmatrix} \frac{d}{ad-b^2} & 0 \\ 0 & \frac{a}{ad-b^2}\end{pmatrix}\\
   \Gamma^{(\nmif, \unit)}_{\nmif, \unit} &= \begin{pmatrix} a - \frac{b^2}{d} & 0 \\ 0 & d-\frac{b^2}{a}\end{pmatrix}, 
\end{align*}
and therefore the precision matrix after the marginalization step is 
\begin{align}
\begin{split}
 \Gamma^{(\phi, u)}_{\nmif, \unit} &= \begin{pmatrix}
    \tau_{\nmif, \unit-1}^{(\phi)} + \Lambda_{\phi}^{(u)} - \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(u)} + \Lambda_u} & 0 \\
    0 & \tau_{\nmif, \unit-1}^{(u)} + \Lambda_u - \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(\phi)} + \Lambda^{(u)}_{\phi}}
   \end{pmatrix} \\
   &= \begin{pmatrix}
    \tau_{\nmif, \unit-1}^{(\phi)} + \Lambda_{\phi}^{(u)} - \alpha_{\nmif, \unit} & 0 \\
    0 & \tau_{\nmif, \unit-1}^{(u)} + \Lambda_u - \beta_{\nmif, \unit}
   \end{pmatrix},
   \end{split}\label{eq:margPrecision}
\end{align}
with $\alpha_{\nmif, \unit} = \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(u)} + \Lambda_u}$ and $\beta_{\nmif, \unit} = \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(\phi)} + \Lambda^{(u)}_{\phi}}$. Because $\Lambda^*_u$ is a positive definite matrix, for all real numbers $c > 0$,

\begin{align*}
    \Lambda_{\phi}^{(u)} > \frac{\Lambda^2_{\phi, u}}{\Lambda_{u}} > \frac{\Lambda^2_{\phi, u}}{c + \Lambda_{u}} 
  \end{align*}
And therefore by letting $\alpha = \min_u \big(\Lambda_u - \frac{\Lambda^2_{\phi, u}}{\Lambda^{(u)}_{\phi}}\big) > 0$, we have $\tau_{\nmif, \unit}^{(\phi)} > \tau_{\nmif, \unit-1} + \alpha$ for all $\nmif \in \mathbb{N}$ and $\unit \in \seq{1}{\Unit}$. 
By iterating this inequality, we have $\tau_{\nmif, \unit}^{(\phi)} > \tau^{(\phi)}_0 + \nmif\alpha$, and thus we see that $\tau_{\nmif, \unit}^{(\phi)} = O(\nmif)$. 
% \eic{should $n$ be $nU$?} 
% \jwc{No, U is constant, so it's the same asymptotic order.}
Therefore as $\nmif \rightarrow \infty$, $\tau_{\nmif, \unit}^{(\phi)} \rightarrow \infty$. 
A similar calculation shows that $\tau_{\nmif, k}^{(\unit)} = O(\nmif)$ for all $k, \unit \in \seq{1}{\Unit}$. 
Thus, considering the covariance matrix after $\nmif$ complete iterations, 
$$
\big\|\Sigma_{\nmif}\big\|_2 = \big\|\Gamma^{-1}_\nmif\big\|_2 = O(1/\nmif),
$$
% \eic{NOTE: $x_n=y_n=)(1/\nmif)$ IS PROBLEMATIC BECAUSE IT REVEALS THE ABUSE OF $=$ COMMON IN $o/O$ NOTATION. REALLY, $O(1/\nmif)$ IS AN EQUIVALENCE CLASS, AND WE SHOULD SAY $\in O(1/\nmif)$. LET'S DISCUSS}
% jwc{Thanks for the recommendation. I was following the definitions / notation used by Keener's textbook, in which case the equality is okay (though maybe abused).}
proving the statement in Theorem~\ref{theorem:GG} that $\|\Sigma_\nmif\|_2 \rightarrow 0$. 
To finish the proof, we need to show that $|\mu_\nmif|_2 \rightarrow 0$. 
To do this, we need more precise descriptions for the rates at which the precision grows.

>>>>

As a reminder, we assume that there are $U\geq 2$ units, labeled $\seq{1}{U}$, with the data for unit $u$ being denoted as $\bm{y}^*_u$. 
We write $\theta = (\phi, \psi_1, \ldots, \psi_U)$, and recall that the unit likelihood $L_{u}(\theta;\bm{y}^*_u)$ for unit $u \in 1:U$ depends only on the shared parameter $\phi$ and the unit-specific parameter $\psi_u$.
The panel assumption implies that, conditioned on the parameter vector, the units are dynamically independent.
Therefore, we can decompose the likelihood function for the entire collection of data $L(\theta; \bm{y}^*)$ as:
\begin{align*}
    L(\theta; \bm{y}^*) = \prod_{u = 1}^U L_u(\theta;\bm{y}_u^*) = \prod_{u = 1}^U L_u(\phi, \psi_u;\bm{y}_u^*).
\end{align*}

@@

We now show that the mean of the resulting Gaussian distribution converges to the MLE, while the covariance matrix converges to the zero matrix, resulting in a density with all mass centered at the MLE. 
To aid this calculation, we introduce the following lemma.
\begin{lemma}\label{lemma:bound}
\label{lemma:matrix}
  Let $d$ be a positive integer, and let $B_k \in \R^{2\times 2}$ for $k \in \seq{1}{d}$ be a collection of real-valued matrices.
  We construct a sequence of matrices $A_{k} \in \R^{d+1\times d+1}$ such that:
  \begin{equation*}
    [A_k]_{i, j} = \begin{cases}
        [B_k]_{1,1}, & i = j = 1 \\ 
        [B_k]_{1,2}, & i = 1, j = k+1 \\
        [B_k]_{2, 1}, & i = k+1, j=1 \\
        [B_k]_{2,2}, & i = j = k+1 \\
        1, & i = j, \, i \notin \{1, k+1\} \\
        0, & \text{otherwise}
	\end{cases}
  \end{equation*}
 That is, $A_k$ is a perturbation of the identity matrix, where the first and $(k+1)$th row and column have been modified on the diagonal and on their off-diagonal intersection to match the matrix $B_k$.
If for all $k \in \seq{1}{d}$, $\|B_k\|_2 \leq c$ for some constant $0 < \lemmaBound \leq 1$, then
\begin{equation}\label{eq:lemma:bound}
\bigg\| \prod_{k = 1}^d A_k \bigg\|_2 \leq \lemmaBound.
\end{equation}
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lemma:bound}]
  For $i, j \in \seq{1}{(d+1)}$, denote $\mu_{(i, j)} \in \mathbb{R}^2$ as the sub-vector of $\mu \in \mathbb{R}^{d+1}$ that contains only the $i$ and $j$th elements, and write $\mu_{-(i, j)} \in \mathbb{R}^{d-1}$ to be the sub-vector of $\mu$ after removing these elements.
  By design, the matrix $A_k$ operates only on the sub-vector $\mu_{(1, k+1)}$.
That is, if $B_k \, \mu_{(1, k+1)} = (\tilde{\mu}_{(1)}, \tilde{\mu}_{(k+1)})^T$, then $A_k\mu = (\tilde{\mu}_{(1)}, \mu_{(2)}, \ldots, \tilde{\mu}_{(k + 1)}, \ldots, \mu_{(d+1)})$.
  From this, we see that for positive integer $m \leq d$,
  the first $m$ factors of the product $\prod_{k = 1}^d A_k$ modify only the first $m + 1$ dimensions of a vector $\mu \in \R^{d+1}$.

We proceed by mathematical induction on the dimension size.
Let $\big\{A^{(d)}_k, k\in \seq{1}{d-1}\big\}$ be a collection of matrices satisfying the condition of Lemma~\ref{lemma:bound} for each $d$.
Setting $P_d=\big\|\prod_{k = 1}^{d-1} A^{(d)}_k \big\|_2$, we first observe that Eq.~(\ref{eq:lemma:bound}) holds for $d=2$ as a direct consequence of the condition $\|B_k\|_2 \leq c$.
Suppose inductively that the lemma holds for $d$, so that $P_d\leq\lemmaBound$.
We wish to bound $P_{d+1}$.
We can choose $A^{(d)}_k$ to be the $(d+1)\times (d+1)$ sub-matrix of $A^{(d+1)}_k$ omitting row and column $(d+2)$.
Thus, for $k\in\seq{1}{d}$,
\begin{equation*}
A^{(d+1)}_k =
  \begin{pmatrix}
    A^{(d)}_k & 0 \\
    0 & 1
  \end{pmatrix}.
\end{equation*}
Consider a vector $\mu \in \R^{d+1}$, such that $\|\mu\|_2 \leq 1$.
Let $\tilde{\mu} \in \R^{d}$ be defined by 
\begin{equation}
\tilde{\mu} = \bigg[\bigg(\prod_{k = 1}^{d-1} A^{(d+1)}_k\bigg)\mu\bigg]_{(1:d)}
\end{equation}
and notice that we have
\begin{equation}\label{eq:lemma:d}
\tilde{\mu} = \bigg(\prod_{k = 1}^{d-1} A^{(d)}_k\bigg)\mu_{(1:d)}.
\end{equation}
By construction, we have 
\begin{align*}
    \bigg|\Big(\prod_{k = 1}^d A^{(d+1)}_k\Big)\mu \bigg|^2_2 &= \bigg|\Big(A^{(d+1)}_d\prod_{k = 1}^{d-1} A^{(d+1)}_k\Big)\mu \bigg|^2_2\\
    &= \big|A^{(d+1)}_d(\tilde{\mu}_{(1:d)}, \mu_{(d+1)})^T\big|^2_2 \\
    &= \big|B^{(d+1)}_d (\tilde{\mu}_1, \mu_{(d+1)})^T\big|^2_2 + \big|\tilde{\mu}_{(1:d)}\big|^2_2 - \tilde{\mu}^2_1.
\end{align*}
Because $\|B^{(d+1)}_d\|_2 \leq \lemmaBound$, $|B^{(d+1)}_d (\tilde{\mu}_1, \mu_{(d+1)})^T|^2_2 \leq \lemmaBound^2|(\tilde{\mu}_1, \mu_{(d+1)})^T|^2_2$.
Furthermore, our inductive hypothesis applied to Eq.~(\ref{eq:lemma:d}) implies that
$|{\tilde{\mu}}_{(1:d)}|^2_2 \leq \lemmaBound^2|\mu_{(1:d)}|^2_2 \leq \lemmaBound^2|\mu|^2_2$.
Therefore,
\begin{align}
    \bigg|\Big(\prod_{k = 1}^d A_k\Big)\mu\bigg|^2_2 &\leq \lemmaBound^2({\tilde{\mu}}^2_1 + \mu_{(d+1)}^2) + \lemmaBound^2(|\mu_{(1:d)}|^2_2) - {\tilde{\mu}}^2_1 \\
    &= (\lemmaBound^2-1){\tilde{\mu}}^2_1 + \lemmaBound^2 |\mu|^2_2
    \\
    \label{eq:lemma:conclusion}
    &\leq \lemmaBound^2 \vert\mu\vert^2_2,
\end{align}
with Eq.~(\ref{eq:lemma:conclusion}) implied by $\lemmaBound \leq 1$, and hence $(\lemmaBound^2-1) \leq 0$.
It follows immediately from Eq.~(\ref{eq:lemma:conclusion}) that $P_{d+1}\le c$, completing the proof.
\end{proof}

We now return to the main argument.

@@

Performing the Bayes update to this distribution (Eq.~\ref{eq:margBayes}) results in an unmarginalized precision matrix, which we denote as $\tilde{\Gamma}_{\nmif, \unit}$, where the only modified components are
$$
\tilde{\Gamma}^{(\nmif, \unit)}_{\nmif, \unit} = \Gamma^{(\nmif, \unit)}_{\nmif, \unit-1} + \Lambda_\unit^* = \begin{pmatrix}
  \tau^{(\phi)}_{\nmif, \unit-1} + \Lambda^{(\unit)}_{\phi} & \Lambda_{\phi, \unit} \\ 
  \Lambda_{\phi, \unit} & \tau^{(\unit)}_{\nmif, \unit-1} + \Lambda_{\unit}
\end{pmatrix}. 
$$
The corresponding mean $\tilde{\mu}_{(\nmif, \unit)} = \mu_{(\nmif, \unit)}$, noting that the marginalization procedure (Eq.~\ref{eq:MPIFupdate}) does not affect the mean, remains the same as the previous iteration except for the components
\begin{align}
\begin{split}
    \tilde{\mu}^{(\phi, \unit)}_{\nmif, \unit} &= \big(\tilde{\Gamma}^{(\nmif, \unit)}_{\nmif, \unit}\big)^{-1}\big(\Gamma^{(\nmif, \unit)}_{\nmif, \unit-1} \mu_{\nmif, \unit-1}^{(\phi, u)} + \Lambda_u^* (0, 0)^T\big) \\
   \mu^{(\phi, u)}_{\nmif, \unit} &= \big(\tilde{\Gamma}^{(\phi, u)}_{\nmif, \unit}\big)^{-1}\Gamma^{(\phi, u)}_{\nmif, \unit-1} \mu_{\nmif, \unit-1}^{(\phi, u)} \\
   &= B_{\nmif, \unit} \mu_{\nmif, \unit-1}^{(\phi, u)}, 
\end{split}\label{eq:submatB}
\end{align}
where $B_{\nmif, \unit} = \big(\tilde{\Gamma}^{(\phi, 1)}_{\nmif, \unit}\big)^{-1}\Gamma^{(\phi, u)}_{\nmif, \unit-1}$ is a $2\times 2$ matrix that provides the update to the first and $(\unit+1)$th components of a vector $\mu \in \mathbb{R}^{\Unit+1}$.
Now by defining matrix $A_{\nmif, \unit}$ to be a perturbation of the $\Unit+1$ identity matrix, such that
\begin{equation*}
\big[A_{\nmif, \unit}\big]_{i, j} = \begin{cases}
  [B_{\nmif, \unit}]_{1, 1} & i = j = 1 \\
  [B_{\nmif, \unit}]_{2, 2} & i = j = u+1 \\
  [B_{\nmif, \unit}]_{1, 2} & i = 1, j = k+1 \\
  [B_{\nmif, \unit}]_{2, 1} & i = k+1, j = 1 \\
  1 & i = j, i \notin \{1, k+1\} \\
  0 & \text{otherwise}
\end{cases},
\end{equation*}
We see that for a given precision matrix $\Gamma_{\nmif, \unit-1}$, the a single sub-iteration of Eqs.~\ref{eq:margBayes} and \ref{eq:MPIFupdate} corresponds to a linear update of the mean vector $\mu_{\nmif, \unit-1}$, defined by
\begin{align*}
  \mu_{\nmif, \unit} = A_{\nmif, \unit} \, \mu_{\nmif, \unit-1}.
\end{align*}
Thus, the mean after $M$ complete iterations, denoted by $\mu_{M}$ is calculated by the product of these linear transformations
\begin{align*}
  \mu_M = \bigg(\prod_{\nmif = 1}^M\prod_{u = 1}^U A_{\nmif, \unit}\bigg)\mu_0. 
\end{align*}
Thus, the long-term value of $\mu_M$ depends on the long-term behavior of the product of matrices $A_{\nmif, \unit}$.

We now consider the limiting behavior of the precision matrices, which determines the behavior of the matrices $A_{\nmif, \unit}$. 
First, we recall that the unmarginalized precision matrix is of the form:
\begin{align*}
  {\tilde{\Gamma}}^{(\nmif, \unit)}_{\nmif, \unit} = \begin{pmatrix}
  \tau^{(\phi)}_{\nmif, \unit-1} + \Lambda^{(u)}_{\phi} & \Lambda_{\phi, u} \\ 
  \Lambda_{\phi, u} & \tau^{(u)}_{\nmif, \unit-1} + \Lambda_{u}\end{pmatrix}. 
\end{align*}
The marginalization step (Eq.~\ref{eq:MPIFupdate}) for Gaussian densities is represented by setting the off-diagonal elements of the covariance matrix to be zero.
Thus, if we generalize this step by writing
$
{\tilde{\Gamma}}^{(\nmif, \unit)}_{\nmif, \unit} = \begin{pmatrix}a & b \\ b & d \end{pmatrix},
$
then
$
\big({\tilde{\Gamma}}^{(\nmif, \unit)}_{\nmif, \unit}\big)^{-1} = \frac{1}{ad - b^2}\begin{pmatrix}d & -b \\ -b & a\end{pmatrix}
$
and the marginalized version of the matrix is
$
\big(\Gamma^{(\nmif, \unit)}_{\nmif, \unit}\big)^{-1} = 
    % \frac{1}{ad - b^2}\begin{pmatrix}d & 0 \\ 0 & a\end{pmatrix}\\ &=
    \begin{pmatrix} \frac{d}{ad-b^2} & 0 \\ 0 & \frac{a}{ad-b^2}\end{pmatrix};
$
by taking the inverse of this matrix, we get the precision matrix after a unit-update to be
$
   \Gamma^{(\nmif, \unit)}_{\nmif, \unit} = \begin{pmatrix} a - \frac{b^2}{d} & 0 \\ 0 & d-\frac{b^2}{a}\end{pmatrix}.
$

Using this general calculation, the precision matrix after the marginalization step is 
\begin{align}
 \Gamma^{(\phi, u)}_{\nmif, \unit} &= \begin{pmatrix}
    \tau_{\nmif, \unit-1}^{(\phi)} + \Lambda_{\phi}^{(u)} - \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(u)} + \Lambda_u} & 0 \\
    0 & \tau_{\nmif, \unit-1}^{(u)} + \Lambda_u - \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(\phi)} + \Lambda^{(u)}_{\phi}}
   \end{pmatrix} \nonumber \\
   &= \begin{pmatrix}
    \tau_{\nmif, \unit-1}^{(\phi)} + \Lambda_{\phi}^{(u)} - \alpha_{\nmif, \unit} & 0 \\
    0 & \tau_{\nmif, \unit-1}^{(u)} + \Lambda_u - \beta_{\nmif, \unit}
   \end{pmatrix},\label{eq:margPrecision}
\end{align}
with $\alpha_{\nmif, \unit} = \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(u)} + \Lambda_u}$ and $\beta_{\nmif, \unit} = \frac{\Lambda_{\phi, u}^2}{\tau_{\nmif, \unit-1}^{(\phi)} + \Lambda^{(u)}_{\phi}}$. Because $\Lambda^*_u$ is a positive definite matrix, for all real numbers $c > 0$, 
\begin{align*}
    \Lambda_{\phi}^{(u)} > \frac{\Lambda^2_{\phi, u}}{\Lambda_{u}} > \frac{\Lambda^2_{\phi, u}}{c + \Lambda_{u}} 
  \end{align*}
And therefore by letting $\alpha = \min_u \big(\Lambda_u - \frac{\Lambda^2_{\phi, u}}{\Lambda^{(u)}_{\phi}}\big) > 0$, we have $\tau_{\nmif, \unit}^{(\phi)} > \tau_{\nmif, \unit-1} + \alpha$ for all $\nmif \in \mathbb{N}$ and $\unit \in \seq{1}{\Unit}$. 
By iterating this inequality, we have $\tau_{\nmif, \unit}^{(\phi)} > \tau^{(\phi)}_0 + \nmif\alpha$, and thus we see that $\tau_{\nmif, \unit}^{(\phi)} = O(\nmif)$. 
Therefore as $\nmif \rightarrow \infty$, $\tau_{\nmif, \unit}^{(\phi)} \rightarrow \infty$. 
A similar calculation shows that $\tau_{\nmif, k}^{(\unit)} = O(\nmif)$ for all $k, \unit \in \seq{1}{\Unit}$. 
Thus, the covariance matrix after $\nmif$ complete iterations has zeros on off-diagonal elements, and the diagonal elements are of order $O(1/\nmif)$, proving the statement in Theorem~\ref{theorem:GG} that $\|\Sigma_\nmif\|_2 \rightarrow 0$. 
To finish the proof, we need to show that $|\mu_\nmif|_2 \rightarrow 0$. 
To do this, we need more precise descriptions for the rates at which the precision grows.

Our previous calculations establish that both $\tau^{(\phi)}_{\nmif, k}$ and $\tau^{(\unit)}_{\nmif, k}$ are strictly increasing sequences in both $\nmif$ and $k$, and are unbounded. 
Thus, the correction terms $\alpha_{\nmif, \unit}$ and $\beta_{\nmif, \unit}$ converge to zero, as the only moving parts are the precision terms that are in the denominator of each of these sequences. 
Furthermore, the sequence $\{\alpha_i\}$, defined as
\begin{align*}
  \alpha_i = \sum_{u = 1}^U \alpha_{i, u} 
\end{align*}
satisfies $\alpha_i \rightarrow 0$. 
Thus, by iterating Eq.~\ref{eq:margPrecision}, we can express the precision matrix after $\nmif$ complete iterations as
\begin{align*}
  \Gamma^{(\phi, u)}_{\nmif} = \Gamma^{(\phi, \unit)}_{\nmif-1, \Unit} = \begin{pmatrix}
    \tau^{(\phi)}_{0} + \nmif\sum_{k = 1}^{\Unit}\Lambda^{(k)}_{\phi} - \sum_{i = 1}^\nmif\alpha_{i} & 0 \\
    0 & \tau^{(\unit)}_{0} + \nmif\Lambda_{k} - \sum_{i = 1}^\nmif \beta_{i, \unit}
  \end{pmatrix}.
\end{align*}
Using the fact that the Ces\`aro mean of a convergent sequence converges to the same limit as the sequence, we have $\frac{1}{\nmif}\sum_{i = 1}^\nmif \alpha_{i} \rightarrow 0$, and therefore 
\begin{align}
  \Gamma^{(\phi, k)}_{\nmif, \unit} &= \begin{pmatrix} 
    \nmif\sum_{k = 1}^\Unit \Lambda_{\phi}^{(k)} + o(\nmif) & 0 \\
    0 & \nmif\Lambda_k + o(\nmif)
  \end{pmatrix}. \label{eq:nGamma}
\end{align}
